---
title: "PA1_template"
author: "Giuseppe Musumeci"
date: "Saturday, December 06, 2014"
output: html_document
---

This is the R Markdown document developed for Reproducible Research: Peer Assessment 1.

##Loading and preprocessing the data

```{r}
data<-read.table("activity.csv",header=TRUE,sep=",",na.strings="NA")
data_aggregate_date<-aggregate(steps ~ date, data, sum)
```
## What is mean total number of steps taken per day?
1.  Plot data using histogram

```{r}
barplot(data_aggregate_date$steps, names.arg = data_aggregate_date$date, xlab = "date", ylab = "steps",col = "lightblue")
```

2.  Calculate and report the mean and median total number of steps taken per day

```{r}
mean(data_aggregate_date$steps)
median(data_aggregate_date$steps)
```

##What is the average daily activity pattern?

1.  Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
data_aggregate_interval<-aggregate(steps ~ interval, data, mean)
plot(data_aggregate_interval, type = "l")
```

2.  Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
data_aggregate_interval[which.max(data_aggregate_interval$steps),1]
```

#Imputing missing values

1.  Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(data))
```
2.  Devise a strategy for filling in all of the missing values in the dataset. 

The strategy is to use the mean tfor that 5-minute interval


3.  Create a new dataset that is equal to the original dataset but with the             missing data filled in

```{r}
data_aggregate_interval_mean<-aggregate(steps ~ interval, data, mean)
dataComplete<-merge(data,data_aggregate_interval_mean,by="interval",suffix=c("",".y"))
valNA<-is.na(dataComplete$steps)
dataComplete$steps[valNA]<-dataComplete$steps.y[valNA]
dataComplete<-dataComplete[,c(1:3)]
```
4.  Make a histogram of the total number of steps taken each day 

```{r}
dataComplete_aggregate_date<-aggregate(steps ~ date, dataComplete, sum)
barplot(dataComplete_aggregate_date$steps, names.arg = dataComplete_aggregate_date$date, xlab = "date", ylab = "steps",col = "lightblue")
```

5.  Calculate and report the mean and median total number of steps taken per day. 

```{r}
mean(dataComplete_aggregate_date$steps)
median(dataComplete_aggregate_date$steps)
```

6.  Do these values differ from the estimates from the first part of the assignment? 

The inpact on mean is obviously null cause we have used mean to estimate missing data;  median tends to become equal to the average.


7.  What is the impact of imputing missing data on the estimates of the total daily number of steps?

We calculate % value
```{r}
S1<-sum(dataComplete_aggregate_date$steps)
S2<-sum(data_aggregate_date$steps)
(S1-S2)/S1*100                                                                      
```

#Are there differences in activity patterns between weekdays and weekends?

1.  Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
S1<-sum(dataComplete_aggregate_date$steps)
S2<-sum(data_aggregate_date$steps)
(S1-S2)/S1*100                                                                      
```


2.  Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
# convert date from string to Date class
dataComplete$date<-as.Date(dataComplete$date,"%Y-%m-%d")

# add a new column indicating day of the week
dataComplete$day <- weekdays(dataComplete$date)

# add a new column called day type and initialize to weekday
dataComplete$day_type <- "weekday"

# If day is Saturday or Sunday, make day_type as weekend
for (i in 1:nrow(dataComplete)){
  if (dataComplete$day[i] == "sabato" || dataComplete$day[i] == "domenica"){
    dataComplete$day_type[i] <- "weekend"
  }
}

# convert day_time from character to factor
dataComplete$day_type <- as.factor(dataComplete$day_type)

# aggregate steps as interval to get average number of steps in an interval across all days
dataCompleted_aggr <- aggregate(steps ~ interval+day_type, dataComplete, mean)

# make the panel plot for weekdays and weekends
library(ggplot2)
qplot(dataCompleted_aggr$interval, dataCompleted_aggr$steps, dataCompleted_aggr, color=dataCompleted_aggr$day_type,geom=c("line","point"),xlab="Number of steps", ylab="Interval")+theme(legend.title=element_blank())

```